SOMETHING ABOUT THE DISTRIBUTED AVERAGING BASED APPROXIMATION IN SENSOR NETWORKS

### 2022-10-10 paper.tex
notation change

% Title.
% ------
\title{SOMETHING ABOUT THE DISTRIBUTED AVERAGING BASED APPROXIMATION IN SENSOR NETWORKS}
%
% Single address.
% ---------------
\name{Matthias Blochberger\(^1\), Filip Elvander\(^2\), Toon, ...\thanks{This research work was carried out at the ESAT Laboratory of KU Leuven, in the frame of the SOUNDS European Training Network. This project has received funding from the European Union's Horizon 2020 research and innovation programme under the Marie Skłodowska-Curie grant agreement No.\,956369.}}
\address{\(^1\)KU Leuven, Dept. of Electrical Engineering (ESAT), STADIUS, 3001 Leuven, Belgium\\\(^2\)Aalto University, Dept. of Signal Processing and Acoustics, 02150 Espoo, Finland}
%
% For example:
% ------------
%\address{School\\
%	Department\\
%	Address}
%
% Two addresses (uncomment and modify for two-address case).
% ----------------------------------------------------------
% \twoauthors{M. Blochberger
%     \thanks{This research work was carried out at the ESAT Laboratory of KU Leuven, in the frame of the SOUNDS European Training Network. This project has received funding from the European Union's Horizon 2020 research and innovation programme under the Marie Skłodowska-Curie grant agreement No.\,956369. This research received funding in part from the Research Foundation - Flanders (FWO) grant 12ZD622N.}}
%         {KU Leuven\\
%         Dept. of Electrical Engineering (ESAT), STADIUS\\
%         % STADIUS\\
%         3001 Leuven, Belgium}
%         {Filip Elvander}
%         {Aalto University\\
%         Dept. of Electronic Systems\\
%         % Section on Artificial Intelligence and Sound\\
%         9220 Aalborg, Finland}
%
\begin{document}
% \ninept
%
\maketitle
%
\begin{abstract}
  Distributed signal-processing algorithms in (wireless) sensor networks often aim to decentralize processing tasks to reduce transmission cost and computational complexity or avoid reliance on a single device for processing. In this contribution, we extend a distributed adaptive algorithm for blind system identification that relies on the knowledge of a stacked network-wide solution vector at each node, the computation of which requires either broadcasting or relaying of node-specific values to all other nodes. The extended algorithm employs a distributed-averaging-based estimation scheme to estimate the network-wide norm value by only using the information provided by neighboring sensor nodes. We introduce a mixing factor between instantaneous and recursion values for adaptivity in a time-varying system. The extension leads to a decreased number of transmissions while maintaining estimation performance close to the optimal case. Simulation results show how the mixing factor and the averaging scheme's number of iterations influence performance.
\end{abstract}
%
\begin{keywords}
multi-channel signal processing, distributed signal processing, wireless sensor networks, blind system identification, distributed averaging
\end{keywords}
%
\section{INTRODUCTION}
\label{sec:intro}

Distributed algorithms have been an active area of research for quite some time, with numerous control, optimization, and signal processing applications.
With the ever-growing number of smart multimedia devices in today's surroundings providing ubiquitous processing and communication capabilities, distributed audio and speech signal processing also found their way into the spotlight.
Algorithms for distributed signal estimation \cite{5483092}, noise control and echo cancellation \cite{9670697}, as well as beamforming \cite{6663655,6329934,MARKOVICHGOLAN20154} amongst others, have been proposed.
The task of distributed single-input-multiple-output (SIMO) blind system identification (BSI) had contributions such as the adaptive cross-relation-based (CR) \cite{yuDistributedBlindSystem2014, liuDistributedBlindIdentification2016}.
Furthermore, we recently introduced an adaptive CR-based algorithm \cite{blochbergerDBSI} using the alternating direction method of multipliers (ADMM) \cite{boydDistributedOptimizationStatistical2011}.
All of the mentioned distributed BSI algorithms rely on information shared between neighboring sensor nodes within the network.
However, the CR-based BSI task necessitates a non-triviality constraint on the full system to be identified (we refer the reader to e.g. \cite{huangAdaptiveMultichannelLeast2002,huangClassFrequencydomainAdaptive2003}), which manifests itself as one or more global variables.
In this case, a global variable is a variable, the computation of which requires information from all nodes within the network.
To overcome the need for the network to be fully connected, \cite{yuDistributedBlindSystem2014, liuDistributedBlindIdentification2016} use an average consensus \cite{xiaoFastLinearIterations2004} approach where a secondary recursion estimates the global variable for each signal frame.
The algorithm in \cite{blochbergerDBSI} relies on node-wise values being relayed throughout the network.
Both approaches introduce additional transmissions of variables between nodes, the number of which, depending on network and neighborhood size, can be substantial and even unfeasable.

@@

\section{DISTRIBUTED ADAPTIVE BLIND SYSTEM IDENTIFICATION USING ONLINE-ADMM}
\label{sec:dbsi}
In \cite{blochbergerDBSI}, we proposed an adaptive SIMO BSI algorithm based on Online-ADMM.
In this section, we will introduce parts which are necessary for the reader to understand the extension.
For the full algorithm, we refer the reader to the previous publication.

The coss-realtion problem 
\begin{align}
  \wf_i^{k+1} &= \underset{\wf_i}{\operatorname{argmin}} \, \mathcal{L}_{\rho} (\wf,\hf^k,\uuf^k)\label{eq:general_consensus_admm:local}\\
  \hf^{k+1} &= \underset{\hf, \|\hf\| = 1}{\operatorname{argmin}}\, \mathcal{L}_{\rho} (\wf^{k+1},\hf,\uuf^k)\label{eq:general_consensus_admm:global}\\
  \uuf_i^{k+1} &= \uuf_i^{k} + \rho \left( \wf_i^{k+1} - \tilde{\hf}_i^{k+1} \right),\label{eq:general_consensus_admm:dual}
\end{align}

% The minimization problem for the local variable \(\wf_i\) \eqref{eq:general_consensus_admm:local} can be solved by various algorithms, in this case however we perform the adaptive update step as
% \begin{equation}
%     \wf_i^{m+1} = \wf_i^{m} - \mu \bm{{V}}_i^m \left( \hat{\aRhof}_i^m \wf_i^m + \uuf_i^m + \rho\left(\wf_i^m - \tilde{\hf}_i^{m}\right)\right),\label{eq:online_admm:local_update}
% \end{equation}
% where \(\mu\), (\(0  < \mu\leq 1\)), is a step size and \(\bm{{V}}_i^m = \left(\hat{\aRhof}_i^m + \rho \I \right)^{-1}\) is the regularized inverse Hessian of the problem.
% The implicit overlap-save matrices in \(\bm{{V}}_i^m\) make this a constrained update step which is costly to compute, so in order to reduce computational complexity, we introduce the approximation
% \begin{equation}
%     \hat{\bm{{V}}}_i^m = \operatorname{diag} \left\{ \operatorname{diag} \left\{ \hat{\aRhof}_i^m \right\} + \rho \mtxb{1} \right\}^{-1},
% \end{equation}
% which is straightforward to compute.
It may be readily verified (cf. \cite{boydDistributedOptimizationStatistical2011}) that the solution to \eqref{eq:general_consensus_admm:global} is given by an update step which computes
\begin{equation}
    \hf_i^{m+1} = \frac{\bar{\hf}_i^{m+1}}{\sqrt{\sum_{j \in \Mset} \|\bar{\hf}_j^{m+1}\|^2}}\label{eq:online_admm:consensus_update}
\end{equation}
for each node \(i \in \Mset\) with the local unnormalized consensus
\begin{equation}
    \bar{\hf}_i^{m+1} = \bar{\wf}_i^{m+1} + \frac{1}{\rho} \bar{\uuf}_i^{m}.
\end{equation}
Here, \(\bar{\wf}_i = \frac{1}{N_i} \sum_{j \in \Tset_i} (\wf_j)_i\) and \(\bar{\uuf}_i = \frac{1}{N_i} \sum_{j \in \Tset_i} (\uuf_j)_i\), with \(N_i = | \Tset_i |\), are neigborhood averages of the channel response estimate and the dual variable of channel \(i\) respectively, computed at node \(i\). The values of all \(\|\bar{\hf}_i^{m+1}\|^2\), \(i \in \Mset\) are transmitted and relayed through the network until all nodes can compute the denominator in \eqref{eq:online_admm:consensus_update}.
% Finally, the update of the dual variables is given by
% \begin{equation}
%     \uuf_i^{m+1} = \uuf_i^{m} + \rho \left( \wf_i^{m+1} - \tilde{\hf}_i^{m+1} \right).\label{eq:online_admm:dual_update}
% \end{equation}

POINT: update \eqref{eq:online_admm:consensus_update} requires knowledge of all \(\|\bar{\hf}_i^{m+1}\|^2\), \(i \in \Mset\) and we introduce a distributed averaging-based approach to tackle this.

@@

% \begin{minipage}[b]{1.0\linewidth}
%   \centering
%   \centerline{\includegraphics[width=8.5cm]{image1}}
% %  \vspace{2.0cm}
%   \centerline{(a) Result 1}\medskip
% \end{minipage}
% %
% \begin{minipage}[b]{.48\linewidth}
%   \centering
%   \centerline{\includegraphics[width=4.0cm]{image3}}
% %  \vspace{1.5cm}
%   \centerline{(b) Results 3}\medskip
% \end{minipage}
% \hfill
% \begin{minipage}[b]{0.48\linewidth}
%   \centering
%   \centerline{\includegraphics[width=4.0cm]{image4}}
% %  \vspace{1.5cm}
%   \centerline{(c) Result 4}\medskip
% \end{minipage}
% %
% \caption{Example of placing a figure with experimental results.}
% \label{fig:res}
% %
% \end{figure}

>>>>

% Title.
% ------
\title{SOMETHING ABOUT THE DISTRIBUTED AVERAGING BASED APPROXIMATION IN SENSOR NETWORKS}
%
% Single address.
% ---------------
\name{Matthias Blochberger\(^1\), Filip Elvander\(^2\), Toon, ...\thanks{This research work was carried out at the ESAT Laboratory of KU Leuven, in the frame of the SOUNDS European Training Network. This project has received funding from the European Union's Horizon 2020 research and innovation programme under the Marie Skłodowska-Curie grant agreement No.\,956369. This research received funding in part from the European Union's Horizon 2020 research and innovation program / ERC Consolidator Grant: SONORA (No.\,773268). This paper reflects only the authors' views and the Union is not liable for any use that may be made of the contained information.}}
\address{\(^1\)KU Leuven, Dept. of Electrical Engineering (ESAT), STADIUS, 3001 Leuven, Belgium\\\(^2\)Aalto University, Dept. of Signal Processing and Acoustics, 02150 Espoo, Finland}
%
% For example:
% ------------
%\address{School\\
%	Department\\
%	Address}
%
% Two addresses (uncomment and modify for two-address case).
% ----------------------------------------------------------
% \twoauthors{M. Blochberger
%     \thanks{This research work was carried out at the ESAT Laboratory of KU Leuven, in the frame of the SOUNDS European Training Network. This project has received funding from the European Union's Horizon 2020 research and innovation programme under the Marie Skłodowska-Curie grant agreement No.\,956369. This research received funding in part from the Research Foundation - Flanders (FWO) grant 12ZD622N.}}
%         {KU Leuven\\
%         Dept. of Electrical Engineering (ESAT), STADIUS\\
%         % STADIUS\\
%         3001 Leuven, Belgium}
%         {Filip Elvander}
%         {Aalto University\\
%         Dept. of Electronic Systems\\
%         % Section on Artificial Intelligence and Sound\\
%         9220 Aalborg, Finland}
%
\begin{document}
% \ninept
%
\maketitle
%
\begin{abstract}
  Distributed signal-processing algorithms in (wireless) sensor networks often aim to decentralize processing tasks to reduce transmission cost and computational complexity or avoid reliance on a single device for processing. In this contribution, we extend a distributed adaptive algorithm for blind system identification that relies on the knowledge of a stacked network-wide solution vector at each node, the computation of which requires either broadcasting or relaying of node-specific values to all other nodes. The extended algorithm employs a distributed-averaging-based estimation scheme to estimate the network-wide norm value by only using the information provided by neighboring sensor nodes. We introduce a mixing factor between instantaneous and recursion values for adaptivity in a time-varying system. The extension leads to a decreased number of transmissions while maintaining estimation performance close to the optimal case. Simulation results show how the mixing factor and the averaging scheme's number of iterations influence performance.
\end{abstract}
%
\begin{keywords}
multi-channel signal processing, distributed signal processing, wireless sensor networks, blind system identification, distributed averaging
\end{keywords}
%
\section{INTRODUCTION}
\label{sec:intro}

Distributed algorithms have been an active area of research for quite some time, with numerous control, optimization, and signal processing applications.
With the ever-growing number of smart multimedia devices in today's surroundings providing ubiquitous processing and communication capabilities, distributed audio and speech signal processing also found their way into the spotlight.
Algorithms for distributed signal estimation \cite{5483092}, noise control and echo cancellation \cite{9670697}, as well as beamforming \cite{6663655,6329934,MARKOVICHGOLAN20154} amongst others, have been proposed.
The task of distributed single-input-multiple-output (SIMO) blind system identification (BSI) had contributions such as the adaptive cross-relation-based (CR) \cite{yuDistributedBlindSystem2014, liuDistributedBlindIdentification2016}.
Furthermore, we recently introduced an adaptive CR-based algorithm \cite{blochbergerDBSI} using the alternating direction method of multipliers (ADMM) \cite{boydDistributedOptimizationStatistical2011}.
All of the mentioned distributed BSI algorithms rely on information shared between neighboring sensor nodes within the network.
However, the CR-based BSI task necessitates a non-triviality constraint on the full system to be identified (we refer the reader to e.g. \cite{huangAdaptiveMultichannelLeast2002,huangClassFrequencydomainAdaptive2003}), which manifests itself as one or more global variables.
In this case, a global variable is a variable, the computation of which requires information from all nodes within the network.
To overcome the need for the network to be fully connected, \cite{yuDistributedBlindSystem2014, liuDistributedBlindIdentification2016} use an average consensus \cite{xiaoFastLinearIterations2004} approach where a secondary recursion estimates the global variable for each signal frame.
The algorithm in \cite{blochbergerDBSI} relies on node-wise values being relayed throughout the network.
Both approaches introduce additional transmissions of variables between nodes, the number of which, depending on network and neighborhood size, can be substantial or even unfeasable.

@@

\section{DISTRIBUTED ADAPTIVE BLIND SYSTEM IDENTIFICATION USING ONLINE-ADMM}
\label{sec:dbsi}
In \cite{blochbergerDBSI}, we proposed an adaptive SIMO BSI algorithm based on Online-ADMM.
In this section, we will introduce parts which are necessary for the reader to understand the extension.
For the full derivation and algorithm, we refer the reader to the cited publication.

The coss-relation problem formulation is used relative information between sensor signals to identify the system, i.e., the acoutic or communication channels.
The solution to this problem is found by the minimization problem
\begin{equation}
  \begin{aligned}
      \hat{\h} = \arg \min_{\h} \quad &\h^\herm \hat{\R} \h \\
      \text{s.t. } \quad &\h^\herm \h = 1.
  \end{aligned}\label{eq:frequency_domain:min_prob}
\end{equation}
where 
\begin{align}
  \w_i^{k+1} &= \underset{\w_i}{\operatorname{argmin}} \, \mathcal{L}_{\rho} (\w,\h^k,\uu^k)\label{eq:general_consensus_admm:local}\\
  \h^{k+1} &= \underset{\h, \|\h\| = 1}{\operatorname{argmin}}\, \mathcal{L}_{\rho} (\w^{k+1},\h,\uu^k)\label{eq:general_consensus_admm:global}\\
  \uu_i^{k+1} &= \uu_i^{k} + \rho \left( \w_i^{k+1} - \tilde{\h}_i^{k+1} \right),\label{eq:general_consensus_admm:dual}
\end{align}

% The minimization problem for the local variable \(\w_i\) \eqref{eq:general_consensus_admm:local} can be solved by various algorithms, in this case however we perform the adaptive update step as
% \begin{equation}
%     \w_i^{m+1} = \w_i^{m} - \mu \bm{{V}}_i^m \left( \hat{\aRhof}_i^m \w_i^m + \uu_i^m + \rho\left(\w_i^m - \tilde{\h}_i^{m}\right)\right),\label{eq:online_admm:local_update}
% \end{equation}
% where \(\mu\), (\(0  < \mu\leq 1\)), is a step size and \(\bm{{V}}_i^m = \left(\hat{\aRhof}_i^m + \rho \I \right)^{-1}\) is the regularized inverse Hessian of the problem.
% The implicit overlap-save matrices in \(\bm{{V}}_i^m\) make this a constrained update step which is costly to compute, so in order to reduce computational complexity, we introduce the approximation
% \begin{equation}
%     \hat{\bm{{V}}}_i^m = \operatorname{diag} \left\{ \operatorname{diag} \left\{ \hat{\aRhof}_i^m \right\} + \rho \mtxb{1} \right\}^{-1},
% \end{equation}
% which is straightforward to compute.
It may be readily verified (cf. \cite{boydDistributedOptimizationStatistical2011}) that the solution to \eqref{eq:general_consensus_admm:global} is given by an update step which computes
\begin{equation}
    \h_i^{m+1} = \frac{\bar{\h}_i^{m+1}}{\sqrt{\sum_{j \in \Mset} \|\bar{\h}_j^{m+1}\|^2}}\label{eq:online_admm:consensus_update}
\end{equation}
for each node \(i \in \Mset\) with the local unnormalized consensus
\begin{equation}
    \bar{\h}_i^{m+1} = \bar{\w}_i^{m+1} + \frac{1}{\rho} \bar{\uu}_i^{m}.
\end{equation}
Here, \(\bar{\w}_i = \frac{1}{N_i} \sum_{j \in \Tset_i} (\w_j)_i\) and \(\bar{\uu}_i = \frac{1}{N_i} \sum_{j \in \Tset_i} (\uu_j)_i\), with \(N_i = | \Tset_i |\), are neigborhood averages of the channel response estimate and the dual variable of channel \(i\) respectively, computed at node \(i\). The values of all \(\|\bar{\h}_i^{m+1}\|^2\), \(i \in \Mset\) are transmitted and relayed through the network until all nodes can compute the denominator in \eqref{eq:online_admm:consensus_update}.
% Finally, the update of the dual variables is given by
% \begin{equation}
%     \uu_i^{m+1} = \uu_i^{m} + \rho \left( \w_i^{m+1} - \tilde{\h}_i^{m+1} \right).\label{eq:online_admm:dual_update}
% \end{equation}

POINT: update \eqref{eq:online_admm:consensus_update} requires knowledge of all \(\|\bar{\h}_i^{m+1}\|^2\), \(i \in \Mset\) and we introduce a distributed averaging-based approach to tackle this.

@@

% \begin{minipage}[b]{1.0\linewidth}
%   \centering
%   \centerline{\includegraphics[width=8.5cm]{image1}}
% %  \vspace{2.0cm}
%   \centerline{(a) Result 1}\medskip
% \end{minipage}
% %
% \begin{minipage}[b]{.48\linewidth}
%   \centering
%   \centerline{\includegraphics[width=4.0cm]{image3}}
% %  \vspace{1.5cm}
%   \centerline{(b) Results 3}\medskip
% \end{minipage}
% \hill
% \begin{minipage}[b]{0.48\linewidth}
%   \centering
%   \centerline{\includegraphics[width=4.0cm]{image4}}
% %  \vspace{1.5cm}
%   \centerline{(c) Result 4}\medskip
% \end{minipage}
% %
% \caption{Example of placing a figure with experimental results.}
% \label{fig:res}
% %
% \end{figure}
